Synthesizing Symmetric Lenses

### 2018-07-10 papers/SymmetricLenses/main.tex
a few tweaks to section 2

The simplest combinator is the identity lens \IdentityLens, which takes as an
argument a regular expression $\BRegex$ and performs an identity mapping on data
matching $\BRegex$. The identity lens fully connects the
corresponding data in the two formats: when one format is
updated, the other format gets the exact same value. In our example,
when a \lstinline{name} is updated in one file, the identify lens ensures it is
also updated in the other file. In this manner, both the \CreateR and \CreateL
functions are the identity function ($\CreateR \App s = s$), and the put
functions merely return the first argument ($\PutROf{s_1}{s_2} = s_2$).

@@

We provide typing rules for each of the simple symmetric combinators, most of
which are syntax directed. However, regular expressions have a number of
equivalences on them, and sometimes we must convert one regular expression type
to another, equivalent, regular expression type. Our typing rules include type
equivalence, so if a lens has type $\Regex \Leftrightarrow \RegexAlt$, then it
is also has type $\Regex' \Leftrightarrow \RegexAlt'$, where $\Regex$ is
equivalent to $\Regex'$, and $\RegexAlt$ is equivalent to $\RegexAlt'$. The
flexibility this type equivalence rule provides is useful when a different
representation of a regular expression would facilitate synthesis.

@@

The expected number of bits required to recover data is a well-known concept in
information theory: it is the \emph{entropy} of the data. Calculating the
entropy of data requires a probability distribution over the space of possible
values for the data. Specifically, given a set $S$ and a probability
distribution over that set $P \OfType S \rightarrow \Reals$, the entropy of $S$
with probability distribution $P$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$.

@@

This infrastructure gives us what we need to find likely satisfying lenses from
a collection of satisfying lenses. All such lenses must synchronize data between
the same two SREs $\Regex$ and $\RegexAlt$. For any such lens $\Lens$, we can
calculate the expected number of bits required to recover a string $t$ in
$\LanguageOf{\RegexAlt}$ from a synchronized string $s$ in
$\LanguageOf{\Regex}$. This expectation is the \emph{conditional entropy of
  $\Regex$ given $\RegexAlt$ and $\Lens$}, formally $\sum_{s \in
  \Regex}\ProbabilityOf{\Regex}{s}*\EntropyOf{\SetOf{t \SuchThat
    \Lens.\PutROf{s}{t} = t}}$. The likelihood we assign to $\Lens$ is the sum
of the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ and the
conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$. This metric
assigns higher cost to lenses where knowing the string present in \Regex
provides little information about what string is present in \RegexAlt. With this
metric, bijections have zero cost -- given a string $s \in \Regex$, we know
exactly what string is present in $\RegexAlt$, as exactly one string is
synchronized to it.

>>>>

The simplest combinator is the identity lens \IdentityLens, which takes as an
argument a regular expression $\BRegex$ and performs an identity mapping on data
matching $\BRegex$. The identity lens fully connects the
corresponding data in the two formats: when one format is
updated, the other format gets the exact same value. In our example,
when a \lstinline{name} is updated in one file, the identity lens ensures it is
also updated in the other file. In this manner, both the \CreateR and \CreateL
functions are the identity function ($\CreateR \App s = s$), and the put
functions merely return the first argument ($\PutROf{s_1}{s_2} = s_2$).

@@

We provide typing rules for each of the simple symmetric combinators, most of
which are syntax directed. However, regular expressions have a number of
equivalences on them, and sometimes we must convert one regular expression type
to another, equivalent, regular expression type. Our typing rules include type
equivalence, so if a lens has type $\BRegex \Leftrightarrow \BRegexAlt$, then it
is also has type $\BRegex' \Leftrightarrow \BRegexAlt'$, where $\BRegex$ is
equivalent to $\BRegex'$, and $\BRegexAlt$ is equivalent to $\BRegexAlt'$. The
flexibility this type equivalence rule provides is useful when a different
representation of a regular expression would facilitate synthesis.

@@

The expected number of bits required to recover data is a well-known concept in
information theory: it is the \emph{entropy} of the data. \saz{Cite Shannon?} Calculating the
entropy of data requires a probability distribution over the space of possible
values for the data. Specifically, given a set $S$ and a probability
distribution over that set $P \OfType S \rightarrow \Reals$, the entropy of $S$
with probability distribution $P$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$.

@@

This infrastructure gives us what we need to find likely satisfying lenses from
a collection of satisfying lenses. All such lenses must synchronize data between
the same two SREs $\Regex$ and $\RegexAlt$. For any such lens $\Lens$, we can
calculate the expected number of bits required to recover a string $t$ in
$\LanguageOf{\RegexAlt}$ from a synchronized string $s$ in
$\LanguageOf{\Regex}$. This expectation is the \emph{conditional entropy of
  $\Regex$ given $\RegexAlt$ and $\Lens$}, formally $\sum_{s \in
  \Regex}\ProbabilityOf{\Regex}{s}*\EntropyOf{\SetOf{t \SuchThat
    \Lens.\PutROf{s}{t} = t}}$. The likelihood we assign to $\Lens$ is the sum
of the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ and the
conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$. This metric
assigns higher cost to lenses where knowing the string present in \Regex
provides little information about what string is present in \RegexAlt. With this
metric, bijections have zero cost---given a string $s \in \Regex$, we know
exactly what string is present in $\RegexAlt$, as exactly one string is
synchronized to it.
